%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Journal Article
% LaTeX Template
% Version 1.3 (9/9/13)
%
% This template has been downloaded from:
% http://www.LaTeXTemplates.com
%
% Original author:
% Frits Wenneker (http://www.howtotex.com)
%
% License:
% CC BY-NC-SA 3.0 (http://creativecommons.org/licenses/by-nc-sa/3.0/)
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%----------------------------------------------------------------------------------------
%	PACKAGES AND OTHER DOCUMENT CONFIGURATIONS
%----------------------------------------------------------------------------------------

\documentclass[twoside]{article}

%\usepackage[sc]{mathpazo} % Use the Palatino font
%\usepackage[T1]{fontenc} % Use 8-bit encoding that has 256 glyphs
%\linespread{1.05} % Line spacing - Palatino needs more space between lines
%\usepackage{microtype} % Slightly tweak font spacing for aesthetics

\usepackage[hmarginratio=1:1,top=32mm,columnsep=20pt]{geometry} % Document margins
\usepackage{multicol} % Used for the two-column layout of the document
\usepackage[hang, small,labelfont=bf,up,textfont=it,up]{caption} % Custom captions under/above floats in tables or figures
\usepackage{booktabs} % Horizontal rules in tables
\usepackage{float} % Required for tables and figures in the multi-column environment - they need to be placed in specific locations with the [H] (e.g. \begin{table}[H])
\usepackage{hyperref} % For hyperlinks in the PDF
\geometry{letterpaper}

\usepackage{lettrine} % The lettrine is the first enlarged letter at the beginning of the text
\usepackage{paralist} % Used for the compactitem environment which makes bullet points with less space between them

\usepackage{abstract} % Allows abstract customization
\renewcommand{\abstractnamefont}{\normalfont\bfseries} % Set the "Abstract" text to bold
\renewcommand{\abstracttextfont}{\normalfont\small\itshape} % Set the abstract itself to small italic text

\usepackage{titlesec} % Allows customization of titles
\renewcommand\thesection{\Roman{section}} % Roman numerals for the sections
\renewcommand\thesubsection{\Roman{subsection}} % Roman numerals for subsections
\titleformat{\section}[block]{\large\scshape\centering}{\thesection.}{1em}{} % Change the look of the section titles
\titleformat{\subsection}[block]{\large}{\thesubsection.}{1em}{} % Change the look of the section titles

\usepackage{fancyhdr} % Headers and footers
\pagestyle{fancy} % All pages have headers and footers
\fancyhead{} % Blank out the default header
\fancyfoot{} % Blank out the default footer
\fancyhead[C]{Polytropes $\bullet$ April 21, 2014 } % Custom header text
\fancyfoot[RO,LE]{\thepage} % Custom footer text

%----------------------------------------------------------------------------------------
%	TITLE SECTION
%----------------------------------------------------------------------------------------

\title{\vspace{-15mm}\fontsize{24pt}{10pt}\selectfont\textbf{Polytropes and
Models of White Dwarf Stars}} % Article title

\author{
\large
\textsc{Erin Conn, Matthew Hurley}\\[2mm] % Your name
\normalsize University of North Carolina at Chapel Hill \\ % Your institution
\vspace{-5mm}
}
\date{}

%----------------------------------------------------------------------------------------

\begin{document}

\maketitle % Insert title

\thispagestyle{fancy} % All pages have headers and footers

%----------------------------------------------------------------------------------------
%	ABSTRACT
%----------------------------------------------------------------------------------------

\begin{abstract}

\noindent 

\end{abstract}

%----------------------------------------------------------------------------------------
%	ARTICLE CONTENTS
%----------------------------------------------------------------------------------------

\begin{multicols}{2} % Two-column layout throughout the main article text

\section{Introduction}

\lettrine[nindent=0em,lines=2]{U}nderstanding stellar mechanics requires the use
of mathematical models of the internal structure of stars. We understand stars
to be nearly spherical collections of hot gas held together by self-gravitation
and holding themselves up against gravitational collapse by fluid and radiation
pressure. Equations from Newtonian gravitational and fluid mechanics allow us to
create models for the how the pressure, density, and temperature of stars varies
with their mass and size, but they are complicated and highly
coupled.\cite{hansen2004} Additionally, as mass and density increase, quantum
and relativistic effects become important.

Prior to the development of advanced electronic computers, solutions to these
equations were difficult to impossible to compute. In 1870, American physicist
Jonathan Homer Lane proposed a simplified model\cite{lane1870} by assuming the
gas pressure depends only on the density of the gas (a \emph{polytropic fluid}),
eliminating any explicit dependence on temperature and decoupling the equations
for pressure, temperature, and density.

    \begin{equation}
        \label{eq:polystate}
        P(r) = K\rho^{1+1/n}
    \end{equation}
    
Later Swiss physicist Robert Emden formalized the model in the dimensionless
differential equation that bears their names, the Lane-Emden equation, whose
derivation can be found in the appendix.

    \begin{equation}
        \label{eq:le}
        \frac{1}{\xi^2}\frac{d}{d\xi}\left(\xi^2\frac{d\theta}{d\xi}\right)+\theta^n=0
    \end{equation}

where \(\xi\) is a dimensionless function of the radius, \(\theta\) is a
dimensionless function relating density and pressure, and n is the
\emph{polytropic index} of the fluid. Solutions to this equation are called
\emph{polytropes}.

While Lane's polytropic simplification seemed unrealistic for most situations
conceivable at the time, the vast simplification of its solutions over more
realistic models proved a generous return on the trade with results that were
still close enough to observed data to be very useful. Closed-form
solutions\cite{leblanc2010} can
be found readily found for polytropic indices \(n=0\) and \(n=1\), and with some
algebraic substitution another one can be found for \(n=5\) (which has an
infinite radius and will not be discussed further), and numerical solutions can
be found using techniques known at the time of Lane's original publication.

In the 20th century new utility was found for this model with the development of
quantum mechanics and the discovery of \emph{white dwarfs}, stellar remnants
composed of extremely dense gas. The electron density in white dwarfs approaches
the density of available quantum energy states; and since the Pauli exclusion
principle disallows multiple electrons from occupying the same quantum
state\cite[p.216]{griffithsqm}, this results in what is called \emph{electron
degeneracy pressure}\cite[pp.217-218]{griffithsqm} which is the main force
opposing the dwarf's own gravity since white dwarfs are no longer actively
undergoing fusion. White dwarfs are therefore composed of nearly fully
\emph{degenerate matter}. Serendipitously, fully degenerate matter has the
following equation of state:\cite[pp.163-166]{hansen2004}

    \begin{equation}
        \label{eq:degenstate}
        P(r)=K\rho(r)^{\gamma}
    \end{equation}

identical in form to eq.~\ref{eq:polystate}, \emph{i.e.}, it is a polytropic
gas!

%------------------------------------------------

\section{Methods}

\begin{compactitem}
\item 
\end{compactitem}

%------------------------------------------------

\section{Results}

We obtained the following parameters for nonrelativistic (\(n=1.5\)) and
relativistic (\(n=3\)) polytropes using a Runge-Kutta scheme in Matlab:

\begin{table}[H]
\caption{Solutions obtained with Runge-Kutta}
\centering
\begin{tabular}{c | c c c}
\toprule
%\multicolumn{2}{c}{Name} \\
%\cmidrule(r){1-2}
\(n\) & \(\xi_f\) & \(\theta'(\xi_f)\) & \(\rho_c\) \\
\midrule
1.5 & 3.6838 & -0.2033 & 5.9907 \\
3 & 6.8968 & -0.0424 & 54.1825 \\
\bottomrule
\end{tabular}
\end{table}


Compared with the parameters outlined in \textit{Stellar
Interiors}\cite{hansen2004} any discrepancies can be attributed to rounding
differences.

\begin{table}[H]
\caption{Parameters for \(n=1.5\) and \(n=3\) polytropes\cite{hansen2004}}
\centering
\begin{tabular}{c | c c c}
\toprule
%\multicolumn{2}{c}{Name} \\
%\cmidrule(r){1-2}
\(n\) & \(\xi_f\) & \(\theta'(\xi_f)\) & \(\rho_c\) \\
\midrule
1.5 & 3.6538 & -0.20330 & 5.991 \\
3 & 6.8969 & -0.04243 & 54.1825 \\
\bottomrule
\end{tabular}
\end{table}

%------------------------------------------------

\section{Discussion}

\end{multicols}
%----------------------------------------------------------------------------------------
%	REFERENCE LIST
%----------------------------------------------------------------------------------------

\bibliographystyle{abbrv}
\bibliography{poly}
%----------------------------------------------------------------------------------------


\end{document}
